### 2016-06-29 sections/intro0.0.tex
V1

%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%Intro
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
\section{Introduction}
% ideas about science of paper will add the them at the end

% maybe a little bit about PAH and other properties of nearby galaxies

% data mining in general?ANN?

>>>>

%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%Intro
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
\section{Introduction} %%( I am not really happy with this introduction but it is the best I can have for now:/ )
%General information about galaxies %(Maybe it is too generic?!!)
Inside each galaxy, there are millions of stars and large amount of dust and gas, which all affect on each others' formation, evolution and extinction.
Galaxies are an active and complicated systems, that are constantly changing.
There are extensive amount of studies about correlations between galaxies' emission and their physical properties, both observationally and theoretically %(Some references here). 
These studies expands our knowledge of the underlying processes in formation and evolution of galaxies every day, but the details of how these process work are still undetermined.

% Nearby galaxies and their importance 
Nearby galaxies play an important role in learning about galaxies' formation and evolution.
These galaxies are close enough to be able to observe various regions inside them in details.
As a result, many studies are devoted to find relations between physical properties of the galaxies in both spatially resolved regions and galaxies as a whole%(Some more references here)
, and many others use these results in analyzing data from high-redshift galaxies. %(Some more references here)
The amount of data that is acquired from nearby galaxies is increasing daily.
Advances in technology helps to both observe and store enormous amount of data from sky.%The sky?! 
Telescopes and surveys gather data from nearby galaxies in variuse wavelengths and from these data we can measure physical properties of galaxies (e.g. stellar mass, star formation rate (SFR), dust mass, gas mass)~\citep[e.g.][]{Calzetti07,Dale09,Eskew12}.

% A little bit about M31 and previous studies in M31.
The Andromeda galaxy (M31), with a distance of $\sim$0.78~Mpc~\citep{McConnachie05}, is the closest spiral galaxy to the Milky Way.
Images of this galaxy provide us a detail view of inside of a spiral galaxy from distance.
M31 data was a target of many telescopes (e.g. {\it Hubble} \citep{Dalcanton12}, \Spitzer\citep{Wener04}, \Herschel\citep{Pilbratt10} space telescopes) and was a subject of many more studies \citep[e.g.][and references therein]{Barmby06, Gordon06, Azimlu11, Sanders12, Dim15, Rahmani16}.

\cite{Barmby06} used data from Spitzer Infrared Array Camera \citep[IRAC;][]{Fazio04} to show evidence of Polycyclic Aromatic Hydrocarbons (PAH)s in interstellar medium (ISM) of M31.
Using the mid-IR data from Multiband Imaging Photometer for Spitzer (MIPS), \cite{Gordon06} studied the morphology of the dust in M31.
\cite{Azimlu11} and \cite{Sanders12} catalogued and studied H {\sc II} regions of M31.
\cite{Draine14, Mattsson14, Viaene14, Smith12} and~\cite{Fritz12} used \Herschel data to study dust in the ISM of the galaxy.
Properties of the current stars in the galaxy were studied by~\cite{Tamm12} and~\cite{Massey07}. %%%And some from PHAT survey
\cite{Rahmani16, Ford13} and \cite{Tabatabaei10} measured the SFR in M31 in order to study star formation laws, and \cite{Dim15} and \cite{Kapala15} used spectroscopic data of the galaxy to study PAHs and other line emissions in the ISM.

Although all these studies measured some properties of M31, or answered specific scientific questions about this galaxy, we still do not have a complete picture of undergoing processes in the galaxy.
The amount of the observed data and measured quantities of M31 makes this galaxy a suitable target for a knowledge discovery and data mining study.
Knowledge discovery and data mining methods are designed to extract hidden information from data and are tested in many astronomical studies.
However, to the best of our knowledge, this study is the first study that uses a data mining method on data from nearby galaxies. %%% ehh

% data mining and clustering in general; when we have so many data and we want to map them
\cite{Ball10} wrote an extensive review of data mining and machine learning, and their usages in astronomy.
A data mining algorithm learns about data from sets of trainings, which can be supervised or unsupervised.
The supervised trainings refer to methods, that use examples of the desired output to learn about input data and are valuable tools to classify data with known target values.
On the contrary, the unsupervised methods train without any prior knowledge of output results. 
They solely work based on the underlying structure of input data.   
The unsupervised methods are very useful tools in knowledge discovery studies on input data, which we do not have any insight of them (or when we want to make sure that we did not miss any valuable information in our previous studies). %%% ehh to after or part.

%SOM
One the most tested methods of data mining in astronomy is the artificial neural networks (ANNs)~\citep[e.g.][and references therein]{Hossein12, Hossein14,Hossein16,Ellison16a}.
ANNs are designed to work in the same way that the neurons work in a human mind.
They are networks of interconnected neurons (nodes), which all of the connections are weighted.
These networks are used to study about nonlinear and complex relations between input and output data.
%These relations can be applied to the new sets of data with similar feature

One of the most well-known unsupervised neural network in astronomy is a Kohonen self-organizing map (also called self-organizing map, or SOM).
SOMs map and visualize a complex and nonlinear high dimension data~\citep{Kohonen82} and show simple geometrical relationships in non-linear high dimensional data~\citep{Kohonen98}.
The result of a SOM is a 1D or 2D network of neurons, which shows the positions of clusters and their relative distance.
Since 1990s, many studies utilized SOMs for object classification and clustering (e.g. classifying quasars' spectra, star/galaxy classifications, gamma-ray bursts clustering and classification of light curves), and photometric redshift estimations~\citep[e.g.][]{Odewahn92, Hernandez94, Murtagh95, Maehoenen95,Scaringi09,Geach12,Fustes13,Meusinger16,Rahmani16b} %%% Why I cannot add submitted in front of the high-z paper

% other Unsupervised methods
The purpose of this project is to have a new insight into nearby galaxies using M31 data with focusing on relations between PAHs and other properties of the galaxy.
To fulfill this purpose, an unsupervised data mining method is the most suitable choice. %%eh  
K-means algorithm, SOMs, and hierarchical clustering are main unsupervised methods that are used in astronomical studies~\citep[e.g.][]{DAbrusco12, Aycha16}. %%add one or two more

For both K-means and SOM algorithms, the user must define the number of clusters, and the algorithms decide how to separate the data into desired number of the clusters.
In the hierarchical clustering method, the user must define dissimilarity between the groups, and the algorithm combines (or divides) existing groups based on their dissimilarity and creates a hierarchical structure. 
Comparing SOMs, K-means and hierarchical clustering shows that in some cases hierarchical clustering method miss-classifies the data~\citep[][and references therein]{Mangiameli96}.
We chose SOM method over K-means due to the fact that SOMs not only clusters data, but also shows similarity and dissimilarity between each cluster.
Therefore, we can cluster our sample data and study the underlying structure of data, simultaneously.

%this project 1 M31 and its extensive data we do not have any SOM in nearby galaxies
In this project we apply SOM algorithm on M31 data, and trained 1D and 2D networks.
Using smaller size networks we study the properties of the clusters and investigate relations between PAHs features with other features in M31.
We create 2D SOMs from subsets of data as well as all available data, which helps us to understand the effect of each input in the position of the clusters in the SOM, which leads to prediction of the unmeasured quantities in a galaxy.
Assuming that all the spiral nearby galaxies have similar properties, we apply 2D trained networks on data from the Pinwheel Galaxy (M101), which is another spiral nearby galaxy. 
If our hypothesis is correct, we should be able to see that regions in M31 and M101 with the same positions in the SOM have the same properties.

In Section $\S$~\ref{Sec: data_SOMN}, we present the sample data from M31 and M101, that we use in this study. 
We describe the SOM method in Section $\S$~\ref{sec: method}. 
The results of the 1D SOM networks and studying PAHs in M31 is presented in Section $\S$~\ref{Sec: 1d_cluster}.
In Section $\S$~\ref{sec: 2d_cluster}, we present the result of 2D SOM networks and how we use them to extract information about other galaxies.
In Section $\S$~\ref{sec: summary}, we summarize our results and discuss potential future work in this subject.

%SOM in Astronomy continued.
% Large spectroscopic surveys have made available integrated spectra of millions of galaxies.
% These integrated spectra combine the light of billions of individual stars and nebulae within a galaxy, and
% finding patterns and common characteristics between galaxies can be a complex task.
% \citet{In12} introduced a new clustering tool based on the SOM method for analyzing these large datasets.
% They used $\sim 60000$ spectra from the Sloan Digital Sky Survey \citep[SDSS;][]{Abazajian09} to test their tool, and created very large SOMs to analyze the type of spectra/objects.
% They also generated SOMs from quasars' spectra in order to find unusual types of spectra. 
% Later, \citet{Meusinger16} used these SOMs and updated data from SDSS and other surveys to find a new class of quasars.
% The other application of SOMs is to find outliers or errors in the data.
% \citet{Fustes13} produced a package based on SOM to classify spectra from the GAIA survey that were previously classified as ``unknown'' by the SDSS pipeline. This package can recognize an astronomical object from instrumental errors, and then classify the object based on its spectrum.

%Compare diff methods and say why we chose SOM!

%supervised and unsupervised network

### 2016-06-29 sections/method0.0.tex
V1

\section{METHOD}
\label{sec: method}
 \subsection{Self Organizing Maps}
 \label{sec: som}

 Kohonen Self organizing map (or self organizing map, SOM) is an unsupervised neural network for mapping and visualizing a complex and non linear high dimension data introduced by~\citep{Kohonen82}. 
 The SOM is a technique that shows a simple geometry relationship of a non-linear high dimension data on a map \citep{Kohonen98}.
 The SOM is a clustering method which reduces the dimension of data to lower dimensions usually 1 or 2D while preserving topological features of the original data.
 Results of SOM contains nodes (usually hexagonal ones) that arranged in 1D or 2D arrays.

 Each nodes may contain one or more samples from input data and distance between nodes represents similarity or dissimilarity of underlying samples. 
 In the way that similar data are closer together in the array and the further nodes go from each other, the more dissimilarity appears between their samples.
 %Sahar_self_notes: maybe "some" is better instead of one or more
 Moreover,a weight vector ``\boldit{W}" with same dimension of the input data associates with each node which will be change during the process and has a key factor in a position of nodes in the map.

 \cite{Geach12} presented the application of the SOM and demonstrate the algorithm of the SOM in detail. In this section we are going to talk briefly about the algorithm of the SOM, how we create our maps and a test model which will help to interpret our results.

 \subsubsection{Algorithm of SOM} 
 \label{sec: algorithm}
     Assuming we have a data set which contains vectors, \boldit{V} $\in \Re^n$ and we want to map them on S1 by S2 map. 
     We start by creating S1 $\times$ S2 empty neurons. 
     The initial arrangement of these neurons depends on a map's topology provided by user.
     Since the topology of the map does not have any affect on the final result, we chose hexagonal topology which is the default topology for SOMs.
     Then, we assign a random weight vector \boldit{W} $\in \Re^n$ to each node.
     The process of creating SOM, happens over series of $N$ iterations. 
     During each iteration the weight vectors might change according to the Kohonen learning rule (equation~\ref{equ: weight adj}). 
      In each iteration:
     \begin{enumerate}
        \item Choose a random vector from our data set.
        \item Calculate the euclidean distance for each node j as  $D_j^2= \sum_{i=0}^{i=n} (V_i - W_i)^2$, and find a neuron with ``$D_{j_{min}}$". This neuron is the winner node and is calling Best Matching Unit (BMU). 
        \item  Compute the radius of the neighbourhood of the BMU to find nodes within this radius. The weight vectors of these nodes will be affected in the next steps. This value is arbitrary and initially can be set to be as high as half of the SOM size and then it decades exponentially over each iteration:
        \begin{equation}
            r^t_{BMU} = r^0_{BMU}e^{(-t/\tau)}
        \end{equation}
        where $\tau$ is a decay constant and usually set to be the same as number of iterations, $N$. $r^0_{BMU}$ and $r^t_{BMU}$ is the radius of the neighbourhood at 0th and $t$th iteration, respectively. 
        \item Change the weight vectors of the BMU and all the nodes within r(t) as:
        \begin{equation}
            \label{equ: weight adj}
            w(t+1)=w(t)+L(t) \times R(t) \times(v(t)-w(t))
        \end{equation}
        where $L(t) = L_0 e^{(-t/\tau)}$ is the learning factor which prevents divergence of the SOM and $R(t)=exp(-\frac{D_j^2}{2r^t_{BMU}})$ is the influence rate. $R(t)$ determines how weight of nodes in the neighbourhood of BMU will change.
        \item  Repeat these steps for $N$ times.
     \end{enumerate}

     In order to create SOM, we used {\tiny MATLAB} neural network toolbox~\citep[ NEWSOM,][]{matlabtolbox}.
     {\tiny MATLAB} code, works in two phases. Phase one is the ``ordering phase". 
     This phase starts with maximum neighbourhood distance (which is initially provided by user), and initial learning factor of 0.9 as a default values. 
     The ordering phase continues for requested number of iterations and during that steps changes the parameters in the way that in the last iteration learning factor is 0.02 and the neighbourhood distance is 1.

@@

     We created different SOM with different sizes, initial neighbourhood, and iteration numbers to find the optimize result for our sample.
     We test our code by varying the number of iterations in ordering phase and find $N = 100$  is optimized for our sample.
     In Section~\ref{sec: test model} we used this toolbox to create SOMs from SED of galaxies and learn how this method works and how we can interpret the results.

 \subsection{Test Model}
 \label{sec: test model}
    Before we start using SOM on our data, we apply this model in a test sample. 
    Our test sample contains SED of 142 galaxies between redshift 0.5 $ \leq$ Z $\leq $ 1, which their physical information such as age, stellar mass and specific star formation are available. 
    \cite{Hossein12} used trained network to classified their spectrum based on SED models from \cite{Kinney96} and \cite{Coleman80}.

    We create various SOM maps with different numbers of neurons and neighbourhood

>>>>

\section{METHOD}
\label{sec: method}
 \subsection{Self Organizing Maps}
 \label{sec: som}
 The SOM is a clustering method which reduces the dimension of data to lower dimensions, while preserving topological features of the original data~\citep{Kohonen98}. 
 The results of the SOM are shown with a map of neurons.
 Each neuron has a fixed position in the map, and may contain one or more samples from input data, \boldit{V} $\in \Re^n$.
 A weight vector,~\boldit{W} $\in \Re^n$, with the same dimension as the input data is associated with each node and will be varied during the training process.
 The process of creating SOM, happens over series of $N$ iterations.
 In each iteration, the algorithm calculates the euclidean distance for each node j as  $D_j^2= \sum_{i=0}^{i=n} (V_i - W_i)^2$, and finds a neuron with ``$D_{j_{min}}$". 
 This neuron is the winner node and is calling Best Matching Unit (BMU). 
 The weight vectors in the neighbourhood of BMU will change according to the Kohonen learning rule (equation~\ref{equ: weight adj}). 
  \begin{equation}
            \label{equ: weight adj}
            w(t+1)=w(t)+L(t) \times R(t) \times(v(t)-w(t))
 \end{equation}
where $L(t) = L_0 e^{(-t/\tau)}$ is the learning factor, which prevents the divergence of the SOM and $R(t)=\exp(-\frac{D_j^2}{2r^t_{BMU}})$ is the influence rate, which determines how the weight of each node will change.
Values for number of neurons, L(t) and R(t) are arbitrary. 
~\cite{Geach12} and \cite{Rahmani16b} demonstrate the algorithm of the SOM in more details.

     In order to create SOM, we used {\sc MATLAB} neural network toolbox~\citep[NNT,][]{matlabtolbox}.
     An SOM in {\sc NNT} can be created by {\sc newsom} library which works in two phases; ``ordering phase" and ``tuning phase" . 
     Phase one is the ``ordering phase". 
     This phase starts with maximum neighbourhood distance, and initial high learning factor usually 0.9 which is provided by user. 
     The ordering phase continues for requested number of iterations and it continues till the learning factor reduces to tuning phase leaning factor and the neighbourhood distance reaches to the number set by the user.

@@

     To present our results, we use {\sc nnt}'s built-in plotting tool.
     Specifically, we combine two of the plots in this tool: a hits map, which shows the number of times each neuron has became the winner (hits), and a distance map, which shows the distance between weight vectors of those neurons.
     In the maps, the purple hexagonal shapes shows the neurons. 
     The distances are shown by the grey cycle colours: the darker the colour, the larger the distance between neurons.
     Neurons with zero hits are left empty.
     In Section~\ref{sec: mock_sample} we used {\sc newsom} to create SOMs from a mock sample to illustrate how this method works and how we can interpret the results.

     One feature of the SOM method is that there is no rule and restriction on number of clusters.
     Users must decide the size of networks based on their data set and their usages of the results .
     There has been few attempts to find a rule for restricting the number of neurons based on the input sample \citep[e.g.][]{Vesanto05}, but none of them are certain. 
     We varied size of SOMs from $1\times2$ to $50\times50$ to choose the most suitable size for our maps, and found that based on the size of our sample, the 2D SOMs to be 10$\times$10. 
     We create the final SOMs with initial values for number of iterations in ordering phase, ordering phase learning factor, tuning phase learning factor, and tuning phase neighbourhood distance of 500, 0.9, 0.02, and 1, respectively. 
     %Also for each grid we created different SOM with different, learning factors, neighbourhood distances, and iteration numbers to find the optimize result for our sample.
     %Based on our data we created our final SOMs with following initial values: number of iteration in ordering phase = 1000; ordering phase learning factor = 0.9; tuning phase learning factor= 0.02; and tuning phase neighbourhood distance to be 1.
     %All the other parameters are default values in {\sc newsom} library and cannot be changed by the users.

\subsection{Mock sample}
\label{sec: mock_sample}

         \begin{figure}
                \centering
                \includegraphics[width=0.5\textwidth]{../images0.01/mock_sample.png}
            \caption{SOM of the mock sample. The axes show the position of the neurons. Hexagonal shapes represent the neurons. The grey cycle colours show the differences between weight of each neuron where white is the minimum differences and black has the maximum one.}
            \label{fig: sample}
        \end{figure}

To show how self-organizing maps work, we created a mock sample, which contains a few regions.
We gave two information to each sample; amount of dust and the SFR.
 Therefore, we have a simple input data with low dimension; each entry only can get either 0 or 1 as high or low SFR, and 0 or 0.5 to show high or low amount of dust for the mock regions.
 We generated an SOM with the size of $10 \times 10$ from the sample, using the method that described in Sec.~\ref{sec: som}.
 Fig. ~\ref{fig: sample} shows the SOM of the mock sample. 
 The axes show the position of the neurons in a $10 \times 10$ network and the hexagonal shapes are the neurons.

Using this method, as expected, we are able to divide the mock sample into 4 distinct groups: regions with high SFR and high amount of dust, regions with low SFR and high amount of dust, regions with high SFR and low amount of dust, and regions low SFR and low amount of dust. 
The plot in the Fig.~\ref{fig: sample}, clearly shows these divisions.
In that plot, the upper part belongs to regions with low amount of dust, while the lower part belongs to the high amount of dust.
The left part of the plot, is where regions with low SFR belong to and the right side is for high SFR regions.
Grey to black colours show the border between regions.
This network is considered as a trained network, and can used to cluster any new data set with similar entries.

Having two regions with exactly the same values in all their quantities in the real world is almost impossible. 
Therefore, one can find a network with high amount of the neurons, which separates the input data completely, and clusters them into separate groups.
However, if the input data has high similarity, the number of neurons must be much higher than the number of input samples to be able to separate the groups from each other. 
Therefore, it is up to the user to decide the similarity or dissimilarity between the input data based on the number of neurons.
